The Small World Network Effect in Software Project Teams \author{Kevin Peterson\\ \small \texttt{pete1968@umn.edu} }

### 2013-11-26 paper/main.tex
Added interview citations.

\textit{Software Project Teams will perform best at moderate levels of $Q$. This performance, measured by the number of interest their projects generate, will increase as the level of $Q$ increases for the developer graph. This will continue up to an optimal $Q$ value, after which performance will begin to degrade. Thus, the performance curve given $Q$ will be U-shaped, and match Uzzi's findings\cite{uzzi2005collaboration}.}

@@

A bipartite graph projection is necessary to further analyze the graph. A bipartite projection involves taking the disjoint node sets $P$ and $C$, and representing the graph as relationships between only one of those sets of nodes. A projection onto the project nodes, or $P$, of figure \ref{fig:example_bipartite_graph} is show in figure \ref{fig:example_bipartite_projection_graph}. Here, projects are connected directly, with edges occurring if one or more developer is shared between the projects. It is important to note that the projection could have been done in terms of the contributors, not the projects. This would have led to a graph with nodes of contributors $C$, each being linked by sharing a common project. This approach is another common way of representing these bipartite graphs\cite{newman2001scientific}, and illustrates the fact that a bipartite projections are lossy compared to the data reprented in the original graph\cite{zhou2007bipartite}.\\

One of the main sources of information loss is multiple connections between projects, or projects that share multiple collaborators. When doing a bipartite projection based on projects $P$, a single link could represent one shared developer or many. This is a source of data loss in the projection. The simple approach is to treat one connection the same as many connections between nodes. This is straightforward but lossy\cite{zhou2007bipartite,grossman1995portion}. A better approach is to assign a \textit{weight} to each edge representing repeated links -- or in our case -- multiple shared collaborators\cite{zha2001bipartite,barrat2004architecture}.

 [scale=.5,auto=left]
  \node[P] (p1) at (-5,3)  {$P_1$};
  \node[P] (p2) at (-3,2)  {$P_2$};
  \node[P] (p3) at (-5,1)  {$P_3$};
  \node[P] (p4) at (-2,1)  {$P_4$};

  \draw ({p1}) -- ({p3});
  \draw ({p1}) -- ({p2});
  \draw ({p1}) -- ({p2});
  \draw ({p2}) -- ({p3});
  \draw ({p3}) -- ({p4});
\end{tikzpicture}
\caption{A $P$ projection of the figure \ref{fig:example_bipartite_graph} bipartite graph}
\label{fig:example_bipartite_projection_graph}
\end{figure}

@@

This metric shows us how closely related, or ``clustered,'' that the graph is. It is also the probability that two nodes will be connected if they share a common neighbor\cite{newman2003properties}. In our context, if project $P_1$ and $P_2$ share a developer with $P_3$, this metric represents the probability that $P_1$ and $P_2$ will share a developer.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{images/freecode-graph.png}
\end{center}
\caption{$C^\Delta$ and Project Popularity}
\label{fig:cc_graph}
\end{figure}

@@

Small-world-ness, or $S^\Delta$, is generally between 1 and 3 for most sub-graphs.

>>>>

\textit{Software Project Teams will perform best at moderate levels of $Q$. This performance, measured by the amount of interest their projects generate, will increase as the level of $Q$ increases for the developer graph. This will continue up to an optimal $Q$ value, after which performance will begin to degrade. Thus, the performance curve given $Q$ will be inverse U-shaped, and match Uzzi's findings\cite{uzzi2005collaboration}.}

@@

A bipartite graph projection is necessary to further analyze the graph. A bipartite projection involves taking the disjoint node sets $P$ and $C$, and representing the graph as relationships between only one of those sets of nodes. A projection onto the contributor nodes, or $C$, of figure \ref{fig:example_bipartite_graph} is show in figure \ref{fig:example_bipartite_projection_graph}. Here, contributors are connected directly, with edges occurring if a two contributors share one or more common projects. It is important to note that the projection could have been done in terms of the projects, not the contributors. This would have led to a graph with nodes of projects $P$, each being linked by sharing a common contributor. This approach is a common way of representing these bipartite graphs\cite{newman2001scientific}, and illustrates the fact that a bipartite projections are lossy compared to the data represented in the original graph\cite{zhou2007bipartite}.\\

One of the main sources of information loss is multiple connections between contributors, or contributors that share multiple projects. When doing a bipartite projection based on contributors $C$, a single link could represent one shared project or many. This is a source of data loss in the projection. The simple approach is to treat one connection the same as many connections between nodes. This is straightforward but lossy\cite{zhou2007bipartite,grossman1995portion}. A better approach is to assign a \textit{weight} to each edge representing repeated links -- or in our case -- multiple shared collaborators\cite{zha2001bipartite,barrat2004architecture}.

\begin{figure}
\centering
\begin{tikzpicture}
\tikzset{C/.style={circle,fill=black!20}}

 [scale=.5,auto=left]
  \node[C] (c1) at (-5,3)  {$C_1$};
  \node[C] (c2) at (-3,2)  {$C_2$};
  \node[C] (c3) at (-5,1)  {$C_3$};
  \node[C] (c4) at (-2,1)  {$C_4$};

  \draw ({c1}) -- ({c3});
  \draw ({c1}) -- ({c2});
  \draw ({c1}) -- ({c2});
  \draw ({c2}) -- ({c3});
  \draw ({c3}) -- ({c4});
\end{tikzpicture}
\caption{A contributor ($C$) projection of the figure~\ref{fig:example_bipartite_graph} bipartite graph}
\label{fig:example_bipartite_projection_graph}
\end{figure}

@@

This metric shows us how closely related, or ``clustered,'' that the graph is. It is also the probability that two nodes will be connected if they share a common neighbor\cite{newman2003properties}. In our context, if contributors $C_1$ and $C_2$ both collaborate with a common contributor w $C_3$, this metric represents the probability that $C_1$ and $C_2$ will collaborate.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{images/freecode-graph.png}
\end{center}
\caption{$Q$ and Project Popularity}
\label{fig:cc_graph}
\end{figure}

@@

Small-world-ness, or $Q$, is generally between 1 and 3 for most sub-graphs.
